Carbon Nucleosynthesis

### 2023-04-06 thesis/tex/main.tex

\begin{titlepage}
   \begin{center}
       \vfill
       \textbf{The Galactic Chemical Evolution of Carbon} \\
        Implications for nucleosynthesis
        \vfill
        Research Thesis
        \vfill
        Presented in partial fulfillment of the requirements for graduation \textit{with research distinction} in Astronomy and Astrophysics in the College of Arts and Sciences of The Ohio State University.
        \vfill
        by 
        \vfill
       {Daniel Alexandra Boyea}\\
       \vfill
       Department of Astronomy\\
       The Ohio State University\\
       \vfill
       April 2023
       \vfill
       Thesis Committee:\\
       David Weinberg, Advisor \\
       Jennifer Johnson \\
       Rolando Vald\'es Aguilar\\
       \vfill

@@

\begin{figure}[htp]
    \includegraphics{subgiants_mgh.pdf}
    \includegraphics{subgiants_mgfe.pdf}
    \caption[APOGEE Subgiants]{The [C/Mg] ratio against [Mg/H] (left) and [Mg/Fe] (right) for the sample of subgiants as in \cite{jack_subgiant}. The left panel uses blue and orange to represent high and low $\alpha$ sequences respectively. On the right, we plot more metal rich ([Mh/H]) stars with darker colors.}
    \label{fig:subgiants}
\end{figure}
\chapter{Nucleosynthesis}

@@

\section{Asymptotic giant branch carbon}

@@

        \caption[AGB carbon yields]{The net fraction carbon yield ($M_\text{C}$, the amount of carbon produced by the star divided by $M$) plotted as a function of mass, $M$. Lighter colors represent higher values of metalicity ([M/H]). Each panel represents a different AGB model. K10 reports stars at MoverH -3, -2, etc. I can reproduce C11 and K10 but not other V13 and KL16 :(((}

    \label{fig:y_agb}
\end{figure*}

@@

    \caption[AGB yields delayed-time-distribution]{
    The Delay Time Distribution for AGB STars}

@@

    \caption[AGB yield metallicity dependence]{On the left, we plot the IMF weighted net fractional carbon yields of each AGB model as a function of metalicity.The right panel plots the normalized net carbon AGB yield as a function of time for a population of stars born at the same time at solar metallicity. We note that for the right panel, the minimum of the DTD for V13 is }

@@

\begin{figure}[htp]
    \centering
    \includegraphics{y_c_cc.pdf}
    \caption[CCSNe carbon yields]{The fist plot shows IMF-weighted net fractional CCSNe yield of
        carbon, $y_C^{cc}$ against metalicity, [M/H], for different studies.
        The right plot instead shows [C/O]$^\text{CC}$, the logarithm of the
        abundance ratio of carbon to oxygen if CCSNe were the only carbon
        source, also as a function of metallicity for the same set of studies.
        The black line represents the carbon yield of the fiducial model,
        $y_C^{CC} = 0.004? (Z/Z_{\sun})^{0.3}$, and becomes dashed over regions
        where we have not tested our model.\citep{NKT13}, \citep{LC18},
        \citep{sukhbold+16}, \citep{WW95}
 }
    \label{fig:y_cc}
\end{figure}

@@

We summarise many of the predictions of stellar models in Fig. \ref{fig:y_cc}. 
\begin{equation}
    {\rm [C/O]^{CC}} = \log_{10}\left( \frac{y_{\rm C}^{\rm CC}}{y_{\rm O}^{\rm CC}}\right) - \log_{10} \left( \frac{Z_{C, \sun }}{Z_{O, \sun }} \right)
\end{equation}
Notice how even these ratios

@@

Chemical abundances reach an equilibrium. At least with simple enough assumptions. The new elements synthesized in stars balance against elements lost to galactic winds and locked in stellar reminants. While our galaxy is likely not in perfect equilibrium or described by a single, homogenous chemical region, we can still use an equilibrium approximation to understand total yields. 
\cite{WAF17}. \cite{james_dwarf}, \cite{james+22}

Here, we assume a simple one-zone chemical evolution model. A one zone model is essentially zero-dimensional---the gas and stars are considered to be homogenously mixed, so we can represent all quantaties as scalars.
We use $M_{X}$ to represent the mass of element X

@@

\begin{figure}
    \centering
    \includegraphics[]{analytic.pdf}
    \caption[Reverse fit yields]{Our reverse fitting method to determine total CCSNe contribution}
\end{figure}

@@

\begin{equation}
\begin{cases}
\text{[Mg/Fe]} >0.12-0.13\text{[Fe/H]}, & \text{[Fe/H]}<0\\
\text{[Mg/Fe]} >0.12, & \text{[Fe/H]}>0\\
\end{cases}
\end{equation}

@@

\begin{figure*}
\label{fig:c_evo}
\includegraphics{evo_tracks.pdf}
\includegraphics{evo_slices.pdf}
\caption[Carbon chemical evolution tracks]{The top set of panes show the time evolution of the gas phase of our
    fiducial model, plotted as tracks through \caah and \caafe where each colored line corresponds to the evolution at that galactic radius. 
The bottom set instead shows the galaxies gas-phase \caah and \caafe trend at 5 different time slices. 
}
\end{figure*}

@@

\begin{figure}
\centering
\includegraphics{oob_agb.pdf}

\caption[AGB GCE Models]{Left: the \caah~ \textbf{median} present-day stellar tracks for four different
AGB models (\cristallo, \karakas, \kl, and \ventura ). The median bin [C/Mg] values from \citet{jack_subgiant} are
plotted as black points with 1-std errors marked by horizontal grey dashes.
Right: same as left but for \caafe. While each AGB yield model may
differ in detail, there is minimal effect on the overall [C/Mg] trends.

While we show the entire dataset in Fig. \ref{fig:subgiants}, we chose to compare the
binned data to our models as black points with dashes representing the 1-sigma
spread, as in e.g. Fig. \ref{fig:agb_sims}. \label{fig:agb_sims}
}
\end{figure}

@@

\caption[Adjusted yield models]{Similar to \ref{fig:agb_sims} except the top plot shows the fiducial model with lower and higher values of $\beta$, the C-CCSNe metallicity dependence. The bottom plot is the same except shows varying AGB fractions.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[]{lateburst_eta.pdf}

\caption[Lateburst and reduced-outflow models]{Same as Fig. \ref{fig:agb_sims} but comparing the fiducial model to the reduced outlow model and a lateburst model. (okay, need to totally change approach...)}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[scale=0.9]{cooh_scatter.pdf}
    \caption[Scatter agreement]{The simulated stars of our model at present day plotted on \caah~and colored such that lighter colors represent stars born at greater galactric radii.}
\end{figure*}

@@

\begin{figure*}
\centering
\includegraphics[]{summary.pdf}
\caption[Gas phase abundances]{Present-day gas phase tracks of the fiducial model in \caah~space. M101 is valid comparison to MW ...(van Dokkum et al. 2014)}
\label{fig:gas_phase}
\end{figure*}

@@

@@

\begin{figure}[htp]
    \includegraphics{coofe_agb_extra.pdf}
    \includegraphics{coofe_agb_extra_0.2.pdf}

>>>>

\begin{titlepage}
   \begin{center}
       \vfill
       \textbf{The Galactic Chemical Evolution of Carbon} \\
        Implications for nucleosynthesis
        \vfill
        Research Thesis
        \vfill
        Presented in partial fulfillment of the requirements for graduation \textit{with research distinction} in Astronomy and Astrophysics in the College of Arts and Sciences of The Ohio State University.
        \vfill
        by 
        \vfill
       {Danielle Alexandra Boyea}\\
       \vfill
       Department of Astronomy\\
       The Ohio State University\\
       \vfill
       April 2023
       \vfill
       Thesis Committee:\\
       David Weinberg, Advisor \\
       Jennifer Johnson \\
       Rolando Vald\'es Aguilar\\
       \vfill

@@

\begin{figure}[htp]
    \includegraphics{subgiants_mgh.pdf}
    \includegraphics{subgiants_mgfe.pdf}
    \caption[APOGEE Subgiants]{The [C/Mg] ratio against [Mg/H] (top) and [Mg/Fe] (bottom) for our sample of APOGEE subgiants. On the top, we plot the high and low $\alpha$ sequences using Eq. \ref{eq:high_alpha} in blue and orange. On the bottom, we show more metal rich ([Mg/H]) stars with lighter colors.}
    \label{fig:subgiants}
\end{figure}
\chapter{Nucleosynthesis}

@@

\section{Asymptotic giant branch carbon}\label{sec:agb}

@@

        \caption[AGB carbon yields]{The net fraction AGB carbon yield ($\tilde{y}$) plotted as a function of initial stellar mass $M$. Lighter colors represent higher values of metalicity ([M/H]). Each panel represents yields from one of our 4 AGB models: C11 \citep{cristallo+11, cristallo+15}, K10 \citep{karakas10}, V13 \citep{ventura+13,ventura+14,ventura+18}, and K16 \citep{KL16, karakas+18}.
            %C11 reports values at metallicities $Z=$ 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02; K10 at $Z=$0.0001, 0.004, 0.008, 0.02; V13 at $Z=$0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04; and K16 at $Z=$0.0028, 0.007, 0.014, 0.03.
        }
            \label{fig:y_agb}
\end{figure*}

@@

    \caption[AGB yields delay time distribution]{
        The normalized delay time distribution (DTD) of AGB carbon, plotted as the net carbon enrichment $M(t)$ normalized to the total carbon enrichment mass $M_{\rm tot}$. Left: Each solid curve corresponds to an AGB yield model at solar metallicity (C11, K10, V13, or K16), and the dashed black line represents the delay time distribution of type Ia supernovae ($\propto t^{-1.1}$). Each model has similar DTD, and most AGB enrichment occurs befor 1 Gyr. Right: each curve represents the C11 AGB model at a different metallicities. Higher metallicities cause carbon enrichment to occur earlier and for the DTD to become non-monotonic.
    }

@@

    \caption[AGB yield metallicity dependence]{The net fractional IMF-weighted AGB carbon yield $\Ycagb$ as a function of metallicity for each of our AGB yield models.
    }

@@

\begin{figure}[htp]
    \centering
    \includegraphics{y_c_cc.pdf}
    \caption[CCSNe carbon yields]{Top: The net fractional IMF-weighted CCSNe yield of
        carbon, $\Ycc$ against metalicity,$\log Z/Z_\odot$, for different studies.
        The right plot instead shows the CCSNe carbon to magnisium abundance ratio, defined in Eq. \ref{eq:c_mg_cc}. The black line represents the carbon yield of the fiducial model,
    $\Ycc = 0.0029 + 0.001 (Z/Z_{\odot})$. The pink line represents the C11 AGB model.    We plot yields from \citet{WW95} as red triangles; \citep{sukhbold+16} as orange diamonds, with larger points representing models with higher explodability; \citet{NKT13} as green stars; and \citep{LC18} as blue circles. \citet{LC18} is the only model including rotation. \citet{NKT13} and \citet{WW95} also report Population III yields, plotted with arrows at the very left.}
    \label{fig:y_cc}
\end{figure}

@@

We summarise many of the predictions of stellar models in Fig. \ref{fig:y_cc}. 
\begin{equation}\label{eq:c_mg_cc}
    {\rm [C/O]^{CC}} = \log_{10}\left( \frac{y_{\rm C}^{\rm CC}}{y_{\rm O}^{\rm CC}}\right) - \log_{10} \left( \frac{Z_{C, \sun }}{Z_{O, \sun }} \right)
\end{equation}
Notice how even these ratios

@@

Chemical abundances reach an equilibrium. At least with simple enough assumptions. The new elements synthesized in stars balance against elements lost to galactic winds and locked in stellar reminants. While our galaxy is likely not in perfect equilibrium or described by a single, homogenous chemical region, we can still use an equilibrium approximation to understand total yields.

\cite{james_dwarf}, \cite{james+22}

Following \citet{WAF17}, we assume a simple one-zone chemical evolution model. A one zone model is essentially zero-dimensional---the gas and stars are considered to be homogenously mixed, so we can represent all quantaties as scalars.

We define $M_{X}$ to be the mass of element X in the gas phase, $\dot{M}_\star$ to be the star fomation rate (in M$_\odot$ yr$^{-1}$).

@@

\begin{figure}
    \centering
    \includegraphics[]{analytic.pdf}
    \caption[Reverse fit yields]{Inferred $\Ycc$ against [Mg/H], assuming equilibrium and 3$\times$C11 AGB yields. Blue points are meadians of each metallicity bin with errorbars as the median standard deviation. The orange curve is our best fit model.}
\end{figure}

@@

\begin{equation}\label{eq:high_alpha}
\begin{cases}
\text{[Mg/Fe]} >0.12-0.13\text{[Fe/H]}, & \text{[Fe/H]}<0\\
\text{[Mg/Fe]} >0.12, & \text{[Fe/H]}>0\\
\end{cases}
\end{equation}

@@

\begin{figure}
\centering
\includegraphics{evo_tracks.pdf}
\includegraphics{evo_slices.pdf}
\caption[Carbon chemical evolution tracks]{Top: the time evolution of the gas phase of our
    fiducial model, plotted as tracks through \caah~and \caafe~where each colored line corresponds to the evolution at that galactic radius. 
    Bottom: snapshots of the gas-phase \caah~and \caafe~trend, each line a different time.
}
\label{fig:c_evo}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{oob_agb.pdf}
\caption[AGB GCE Models]{Left: the \caah~median present-day stellar tracks for four different AGB models (C11, K10, V13, K16) with a metallicity-independent $\Ycc$. Right: same except [C/Mg] against [Mg/Fe] for the bin $-0.15<\text{[Mg/H]}<-0.05$. The median bin [C/Mg] values from \citet{jack_subgiant} are plotted as black points with standard median deviations plotted as grey dashes.
 While each AGB yield model may differ in detail, there is minimal effect on the overall [C/Mg] trends.
}
\label{fig:agb_sims}
\end{figure}

@@

\caption[Adjusted yield models]{Similar to \ref{fig:agb_sims} except the top plot shows the fiducial model with lower and higher values of $\zeta$, the C-CCSNe metallicity dependence. The bottom plot is the same except shows models with $f_{\rm AGB}=$0.1, 0.2, and 0.4. Both the $f_{\rm AGB}$ and $\zeta$ influence \caah, but only $f_{\rm AGB}$ has a significant impact on \caafe.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{lateburst_eta.pdf}
\includegraphics{eta.pdf}

\caption[Lateburst and reduced-outflow models]{Same as Fig. \ref{fig:agb_sims} but comparing the fiducial model to two lateburst models (top), and two reduced outflow models (bottom).}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[scale=0.9]{cooh_scatter.pdf}
    \caption[Scatter agreement]{The simulated stars of our model at present day plotted on \caah~and colored such that lighter colors represent stars born at greater galactric radii. Our model is able to explain the scatter and shape, relying on a combination of artificial measurment errors and radial migration.}
\end{figure*}

@@

\begin{figure}
    \includegraphics{nitrogen.pdf}
    \includegraphics{c_n.pdf}
    \caption{Nitrogen almost works?}
\end{figure}

\begin{figure*}
\centering
\includegraphics[]{summary.pdf}
\caption[Gas phase abundances]{Present-day gas phase tracks of the fiducial model in \caah~space. M101 is valid comparison to MW ...(van Dokkum et al. 2014)

    \citep{cooke+17}, \citet{nissen+14}, \citet{fabbian+09}, \citet{berg+19}, \citet{esteban+02}, \citet{esteban+09}, \citet{esteban+14}, \citet{esteban+19},
}
\label{fig:gas_phase}
\end{figure*}

@@

\begin{figure}
    \centering
    \includegraphics{logg_jack.pdf}
\end{figure}

@@

\begin{figure}[htp]
    \centering
    \includegraphics{agb2.pdf}
